### 2016-03-07 Numerical solution of DEs/Chapter1/homework1/LaTeX/homework1.tex
FEIP

\documentclass[12pt]{article}
\usepackage{listings}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amsthm,amssymb}
\lstset{
	numbers=left,
    framexleftmargin=10mm,
    frame=none,
    backgroundcolor=\color[RGB]{245,245,244},
	keywordstyle=\bf\color{blue},
	identifierstyle=\bf,
	numberstyle=\color[RGB]{0,192,192},
	commentstyle=\it\color[RGB]{0,96,96},
	stringstyle=\rmfamily\slshape\color[RGB]{128,0,0},
	showstringspaces=false
    }

\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}

\newenvironment{theorem}[2][Theorem]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
\newenvironment{lemma}[2][Lemma]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
\newenvironment{exercise}[2][Exercise]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
\newenvironment{problem}[2][Problem]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
\newenvironment{question}[2][Question]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
\newenvironment{corollary}[2][Corollary]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}

\begin{document}

\title{Homework 2016-03-02}
\author{Chuan Lu\\ 
13300180056}

\maketitle

\begin{problem}{1}
\text{ }\\
\end{problem}

\begin{proof}
The code is shown as follows.

\begin{lstlisting}[language={MATLAB}]
% homework1
% author: chuanlu
% 2016-03-02
format long
xx = 10 .^ [-1:-1:-16];
yy = (1 - cos(xx)) ./ (xx .^ 2);
semilogx(xx, yy);    
\end{lstlisting}

The result is shown as follows.
\end{proof}

\begin{problem}{2}
\text{ }\\
$(1)$ Prove: If $[a] \in \Z_{n}$ is a unit, then $[a]$ is not a zero divisor.\\
$(2)$ Prove: If $[b] \in \Z_{n}$ is a zero divisor, then $[b]$ is not a unit.
\end{problem}

\begin{proof}

\end{proof}

\begin{problem}{3}
\text{ }\\
Show that every nonzero element of $\Z_{n}$ is either a unit or a zero divisor.
\end{problem}

\begin{proof}

\end{proof}

\begin{problem}{4}
\text{ }\\
Suppose that $[a]$ is a unit in $\Z_{n}$ and $[b]$ is an element of $\Z_{n}$. Prove that the equation $[a]x = b$ has exactly one solution in $\Z_{n}$
\end{problem}

\begin{proof}

\end{proof}

\begin{problem}{5}
\text{ }\\
Suppose that $[a]$ and $[b]$ are both units in $\Z_{n}$.  Show that the product $[a] \cdot [b]$ is also a unit in $\Z_{n}.$ (Note that this confirms closure under multiplication in the group $U_{n})$.
\end{problem}

\begin{proof}

\end{proof}

\begin{problem}{6}
\text{ }\\
Which of the following are Groups? Which of the following are not groups, and why?\\

\indent (1)  $G = \{{2, 4, 6, 8}\}$ in $\Z_{10}$.  Where $a \star b = ab$\\
\indent (2)  $G = \Q^{\ast}$, where $a \star b = \frac{a}{b}$\\
\indent (3) $G = \Z$, where $a \star b = a - b$\\
\indent (4) $G = \{ {2^{x}\mid x \in \Q} \}$, where $a \star b = ab$\\

\end{problem}

\begin{proof}

\end{proof}

\begin{problem}{7}
\text{ }\\
Consider the set $Q =$ \{ $\pm$1, $\pm$i, $\pm$j, $\pm$k\} of the complex matrices as follows:\\
\[
1=
  \begin{bmatrix}
    1 & 0\\
    0 & 1\\
  \end{bmatrix}
\]

\[
i=
  \begin{bmatrix}
    i & 0\\
    0 & $-$i\\
  \end{bmatrix}
\]

\[
j=
  \begin{bmatrix}
    0 & 1\\
    $-$1 & 0\\
  \end{bmatrix}
\]

\[
k=
  \begin{bmatrix}
    0 & i\\
    i & 0\\
  \end{bmatrix}
\]
Show that $Q$ is a group under matrix multiplication by writing out its multiplicaiton table. (Note: $Q$ is called the quartenion group).
\end{problem}
\begin{proof}

\end{proof}

\end{document}