### 2016-09-21 Config Files/Assignment-TEX/assignment.tex
modification

%!TEX program = xelatex

% \documentclass[10pt]{article}
\input{structure.tex}
\usepackage{epstopdf}
\usepackage{graphics}
\usepackage{subfig}
\usepackage{listings}
\lstset{
  numbers=left,
    framexleftmargin=10mm,
    frame=none,
    backgroundcolor=\color[RGB]{245,245,244},
  keywordstyle=\bf\color{blue},
  identifierstyle=\bf,
  numberstyle=\color[RGB]{0,192,192},
  commentstyle=\it\color[RGB]{0,96,96},
  stringstyle=\rmfamily\slshape\color[RGB]{128,0,0},
  showstringspaces=false,
  extendedchars=false
    }
\DeclareGraphicsExtensions{.eps,.ps,.jpg,.bmp}

\begin{document}

\title{Homework 1}{16.9.21, Chuan Lu}

\problem{1}{Rewrite the target function of K-means in matrix form.}
\solution{Solution}{
$f(x) = argmin_{\mathcal{S}}\sum_{i=1}^{k}\sum_{x_j\in S_j}\sum_{1\le m\le p}\lvert x_{jl} - x_{jm} \rvert^2.$
}

\problem{2}{Cluster $Data_i.csv$ with K-means, judge the number of clusters, and compare differences between different evaluating methods.}
\solution{Solution}{
The code of K-means, evaluating and test scripts can be found from attachments(kmeans.R, evaluate.R, hw1.2.R).
}
\solution{Result}{
For Data1.csv, we cluster with k = 3. The result is shown as below.
\begin{figure}[htbp]
\centering
\includegraphics[width = 15cm]{pics/hw2_data1_cluster.eps}
\caption{The clusters of Data1.csv with k = 3}
\label{clusters of data1}
\end{figure}
\\[1mm]

For choosing k, the result of Data1.csv is shown below; The first is the result of Calinski-Harabasz method, the second Hartigan method and the last Gap Statistic. \\[3pt]
The result of CH method is 9, if we may add a limit that $k\le 10.$ The result of H method is 3, and result of GAP statistic is 3.\\[3pt]

\begin{figure}[htbp]
\centering
\subfloat[CH method]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw2_data1_k_CH.eps}}\hfill
\subfloat[H method]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw2_data1_k_H.eps}}\hfill
\subfloat[GAP statistic]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw2_data1_k_GAP.eps}}\hfill
\caption{The three evaluating methods of choosing k for Data1.csv}
\label{EVA Data1}
\end{figure}

For Data2.csv, since each data point is of 3 dims, we cannot plot its clustering result here. But after deploying the evaluating methods we consider k = 3. The type of datas can be find in attachments(CLUSTER\_DATA2.txt). \\[3pt]

The result of CH method is 8, if we limit that $k\le 10$. The result of H method and result of GAP statistic are 3. \\[3pt]
The pics are shown below.

\begin{figure}[htbp]
\centering
\subfloat[CH method]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw2_data2_k_CH.eps}}\hfill
\subfloat[H method]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw2_data2_k_H.eps}}\hfill
\subfloat[GAP statistic]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw2_data2_k_GAP.eps}}\hfill
\caption{The three evaluating methods of choosing k for Data2.csv}
\label{EVA Data2}
\end{figure}

For Data3.csv, we fix k = 3 after evaluating k with H method and GAP statistic; The result of clustering can be found in attachments(CLUSTER\_DATA3.txt). \\[3pt]

The result of CH method is 9, if we limit that $k\le 10$. The result of H method and result of GAP statistic are 3. \\[3pt]
The pics are shown below.

\begin{figure}[htbp]
\centering
\subfloat[CH method]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw2_data3_k_CH.eps}}\hfill
\subfloat[H method]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw2_data3_k_H.eps}}\hfill
\subfloat[GAP statistic]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw2_data3_k_GAP.eps}}\hfill
\caption{The three evaluating methods of choosing k for Data3.csv}
\label{EVA Data2}
\end{figure}

For Data4.csv, we found k = 3 after evaluating k with H method and GAP statistic; The result of clustering can be found in attachments(CLUSTER\_DATA4.txt). \\[3pt]

The result of CH method is 7, if we limit that $k\le 10$. The result of H method and result of GAP statistic are 3. \\[3pt]
The pics are shown below.

\begin{figure}[htbp]
\centering
\subfloat[CH method]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw2_data4_k_CH.eps}}\hfill
\subfloat[H method]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw2_data4_k_H.eps}}\hfill
\subfloat[GAP statistic]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw2_data4_k_GAP.eps}}\hfill
\caption{The three evaluating methods of choosing k for Data4.csv}
\label{EVA Data2}
\end{figure}
}
\solution{Analysis}{
It is suprising that result of Calinski-Harabasz method does not agree with the other methods. I consider it as the result that I misunderstood the meaning of $W(k), B(k)$. I calculate the former as the target function of cluster; the latter as $\sum_{c_i, c_j \in C}\lVert c_i - c_j \rVert^2$, in which $C$ is the set of cluster centers.
}

\problem{3}{Use hierarchical clustering methods to cluster $Data_i.csv$}
\solution{Solution}{
The code of hierarchical clustering and the test script can be found from attachments(hierarchical\_clustering.R, hw1.3.R).
}
\solution{Result}{
It is somehow embarrassing that I still did't know how to draw trees in R without $hcluster()$. The result can be shown is the attachments, (HC\_DATA1.txt and HC\_DATA2.txt). \\[3pt]
Each result is a (n-1) * narg matrix, in which n is the number of data points and narg is the number of dims in each data point. Each row in the matrix(for example, A[1, ]) means a step when two points merged with each other; A[1, 1] and A[1, 2] merged into a new A[1, 1], and the point represented by A[1, 2] is abandoned. \\[3pt]
It is possible to draw the tree and do Tree-Cut with the results.
}

\end{document}