### 2016-10-09 Numerical Algorithm and Case Studies/Homwwork2/assignment.tex
hw

%!TEX program = xelatex

% \documentclass[10pt]{article}
\input{structure.tex}
\usepackage{epstopdf}
\usepackage{graphics}
\usepackage{subfig}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{courier}
\renewcommand{\sfdefault}{phv}

\DeclareGraphicsExtensions{.eps,.ps,.jpg,.bmp}

\begin{document}

\title{Assignment 2}{16.10.2}
\tiny{Code of this assignment are in the attachment, and they can also be found in GitHub: https://github.com/Orcuslc/Learning/tree/master/Numerical\%20Algorithm\%20and\%20Case\%20Studies/Homework2/code}

% \parbox{0.3\textwidth}{
% Chuan Lu}
% \hfill
% \parbox{0.3\textwidth}{
% 13300180056}
% \hfill
% \parbox{0.3\textwidth}{
% chuanlu13@fudan.edu.cn}

\problem{1 \& 2}{}
\solution{Abstract}{
\\[5pt]
Details of the implementation of LU decomposition in MATLAB.
}
\solution{Introduction}{
\\[5pt]
In most cases pivoting is necessary. If we randomly choose a matrix $M\in \mathbb{R}^{n\times n}$, the probability that $M(1, 1) = max_{i}M(i, 1)$ is $\frac{1}{n}$. So the efficiency of pivoting is a great concern. \\[5pt]
In the experiment below, we generate the matrix $M$ by using MATLAB function \texttt{randn(n, n)}; and we use seperately
$$\texttt{[L, U] = lu(A)}$$
$$\texttt{[L, U, P] = lu(A)}$$
for the two problems.
}
\solution{Results \& Discussion}{
After trying for several times with different $n$, the dimension of the random matrix, we found that in most cases, calling \texttt{[L, U] = lu(A)} leads to $L$ not being a lower triangular matrix, but $U$ is always an upper triangular matrix. \\[5pt]
When we count the number of non-zero elements of each row in $L$, we can find out the list of the numbers is a permutation of \texttt{[1, 2, ..., n]}. \\[5pt]
Thus I tried to call \texttt{[L1, U1, P] = lu(A)}, and the $P$ is exactly the permutation matrix: $$PL1 = L$$.
So it is obvious how MATLAB express the $P$: it simply multiplied it with the L; In order to represent this, I guess that MATLAB records the number of lines that need interchanging in each step; when the algorithm ends, it conbines the records to gain a permutation matrix. \\[5pt]
In order to realize $P*b$, we should use the pernutation record: in each of the $n$ steps, change the elements of two lines, and it will only need $O(n)$ time, and no extra space.
}

\problem{3}{}
\solution{Abstract \& Introduction}{
As is known to all, using Cholesky decomposition for a symmetric positive definite matrix will need half of the flops comparing LU decomposition; In the mean time, since there is no need of pivoting for a symmetric positive definite matrix, in each step $n-1$ times of comparing and $2n$ times of exchanging will be saved on average.
}
\solution{Result}{
We generate the symmetric matrix A by
$$\texttt{A = randn(n, n); A = A + A';}$$
and we did the experiment with $n = 4, 6, 8, 10$. The result of calling 
$$\texttt{L = chol(A)}$$
is:
$$\texttt{Error using \textbf{chol}: Matrix must be positive definite}.$$
}
\solution{Analysis}{
If $A\in \mathbb{R}^{n\times n}$ is not positive definite, then according to the properties of symmetric elementary transformations, $PAP^\top$ is still not positive definite, where $P$ is a permutation matrix. \\[5pt]
Thus, we can assume $A(1:2, 1:2)$ is not positive definite. According to the algorithm, we have
$$l_{11} = \sqrt{a_{11}},$$
$$l_{21} = \frac{a_{21}}{l_{11}},$$
$$l_{22} = (a_{22} - l_{21}^2)^{1/2}.$$
That means $l_{22} = a_{22} - \frac{a_{21}^2}{a_{11}} = \frac{a_{22}a_{11} - a_{21}^2}{a_{11}}$. Since A is symmetric and not positive definite, we have $a_{22}a_{11}-a_{21}^2 = a_{22}a_{11}-a{21}a_{12}<0$. So $l_{22}$ does not exist.
}

\problem{4}{}
\solution{Introduction}{
The $LDL^\top$ decomposition is used for Hermitian indefinite matrices.
}
\solution{Discussion}{
I found an introduction to $LDL^\top$ decomposition from Golub's \textit{Matrix Computations}, Page 186. It said the method was developed by Bunch and Parlett. I searched their paper in Google Scholar, but it seemed that our school has no access to siam.org, so I cannot read the paper.
}

\refgroup{
\reference{Gene H. Golub \textit{et al}, \textit{Matrix Computations}, Page 186-187.}
}

\end{document}