### 2016-10-09 Statistical Learning/Homework2/assignment.tex

%!TEX program = xelatex

% \documentclass[10pt]{article}
\input{structure.tex}
\usepackage{epstopdf}
\usepackage{graphics}
\usepackage{subfig}
\usepackage{listings}
\lstset{
  numbers=left,
    framexleftmargin=10mm,
    frame=none,
    backgroundcolor=\color[RGB]{245,245,244},
  keywordstyle=\bf\color{blue},
  identifierstyle=\bf,
  numberstyle=\color[RGB]{0,192,192},
  commentstyle=\it\color[RGB]{0,96,96},
  stringstyle=\rmfamily\slshape\color[RGB]{128,0,0},
  showstringspaces=false,
  extendedchars=false
    }
\DeclareGraphicsExtensions{.eps,.ps,.jpg,.bmp}

\begin{document}

\title{Homework 2}{16.10.9}

% \parbox{0.3\textwidth}{
% Chuan Lu}
% \hfill
% \parbox{0.3\textwidth}{
% 13300180056}
% \hfill
% \parbox{0.3\textwidth}{
% chuanlu13@fudan.edu.cn}

\problem{1}{Explain the relationship between spectral clustering, normalized spectral clustering and graph cut.}
\solution{Solution}{
We assume $k=2$ in the following explanation. \n
For spectral clustering, we already know that the eigenvector $u_2$ corresponding to the second smallest eigenvalue $\lambda_2$ of the Laplacian matrix of the graph is the solution for the minimization
$$ \min_{\tbf{f}} \ \tbf{f}^\top\tbf{L}\tbf{f},$$
$$s.t. \quad \tbf{f}^\top \tbf{f} = 1, \tbf{f}^\top\tbf{1} = 0$$
In fact, if we define 
$$G(f) = G(f_1, \cdots)$$
For Min Cut, if we define $\tbf{f} = (f_1, f_2\cdots f_n)\tr$
}

% \problem{2}{Cluster $Data_i.csv$ with K-means, judge the number of clusters, and compare differences between different evaluating methods.}
% \solution{Solution}{
% The code of K-means, evaluating and test scripts can be found from attachments(kmeans.R, evaluate.R, hw1.2.R).
% }
% \solution{Result}{
% For Data1.csv, we cluster with k = 3. The result is shown as below.
% \begin{figure}[htbp]
% \centering
% \includegraphics[width = 15cm]{pics/hw2_data1_cluster.eps}
% \caption{The clusters of Data1.csv with k = 3}
% \label{clusters of data1}
% \end{figure}
% \\[1mm]

% For choosing k, the result of Data1.csv is shown below; The first is the result of Calinski-Harabasz method, the second Hartigan method and the last Gap Statistic. \\[3pt]
% The result of CH method is 9, if we may add a limit that $k\le 10.$ The result of H method is 3, and result of GAP statistic is 3.\\[3pt]

% \begin{figure}[htbp]
% \centering
% \subfloat[CH method]{ %
% 	\includegraphics[width = 0.3\textwidth]{pics/hw2_data1_k_CH.eps}}\hfill
% \subfloat[H method]{ %
% 	\includegraphics[width = 0.3\textwidth]{pics/hw2_data1_k_H.eps}}\hfill
% \subfloat[GAP statistic]{ %
% 	\includegraphics[width = 0.3\textwidth]{pics/hw2_data1_k_GAP.eps}}\hfill
% \caption{The three evaluating methods of choosing k for Data1.csv}
% \label{EVA Data1}
% \end{figure}

% For Data2.csv, since each data point is of 3 dims, we cannot plot its clustering result here. But after deploying the evaluating methods we consider k = 3. The type of datas can be find in attachments(CLUSTER\_DATA2.txt). \\[3pt]

% The result of CH method is 8, if we limit that $k\le 10$. The result of H method and result of GAP statistic are 3. \\[3pt]
% The pics are shown below.

% \begin{figure}[htbp]
% \centering
% \subfloat[CH method]{ %
% 	\includegraphics[width = 0.3\textwidth]{pics/hw2_data2_k_CH.eps}}\hfill
% \subfloat[H method]{ %
% 	\includegraphics[width = 0.3\textwidth]{pics/hw2_data2_k_H.eps}}\hfill
% \subfloat[GAP statistic]{ %
% 	\includegraphics[width = 0.3\textwidth]{pics/hw2_data2_k_GAP.eps}}\hfill
% \caption{The three evaluating methods of choosing k for Data2.csv}
% \label{EVA Data2}
% \end{figure}

% For Data3.csv, we fix k = 3 after evaluating k with H method and GAP statistic; The result of clustering can be found in attachments(CLUSTER\_DATA3.txt). \\[3pt]

% The result of CH method is 9, if we limit that $k\le 10$. The result of H method and result of GAP statistic are 3. \\[3pt]
% The pics are shown below.

% \begin{figure}[htbp]
% \centering
% \subfloat[CH method]{ %
% 	\includegraphics[width = 0.3\textwidth]{pics/hw2_data3_k_CH.eps}}\hfill
% \subfloat[H method]{ %
% 	\includegraphics[width = 0.3\textwidth]{pics/hw2_data3_k_H.eps}}\hfill
% \subfloat[GAP statistic]{ %
% 	\includegraphics[width = 0.3\textwidth]{pics/hw2_data3_k_GAP.eps}}\hfill
% \caption{The three evaluating methods of choosing k for Data3.csv}
% \label{EVA Data2}
% \end{figure}

% For Data4.csv, we found k = 3 after evaluating k with H method and GAP statistic; The result of clustering can be found in attachments(CLUSTER\_DATA4.txt). \\[3pt]

% The result of CH method is 7, if we limit that $k\le 10$. The result of H method and result of GAP statistic are 3. \\[3pt]
% The pics are shown below.

% \begin{figure}[htbp]
% \centering
% \subfloat[CH method]{ %
% 	\includegraphics[width = 0.3\textwidth]{pics/hw2_data4_k_CH.eps}}\hfill
% \subfloat[H method]{ %
% 	\includegraphics[width = 0.3\textwidth]{pics/hw2_data4_k_H.eps}}\hfill
% \subfloat[GAP statistic]{ %
% 	\includegraphics[width = 0.3\textwidth]{pics/hw2_data4_k_GAP.eps}}\hfill
% \caption{The three evaluating methods of choosing k for Data4.csv}
% \label{EVA Data2}
% \end{figure}
% }
% \solution{Analysis}{
% It is suprising that result of Calinski-Harabasz method does not agree with the other methods. I consider it as the result that I misunderstood the meaning of $W(k), B(k)$. I calculate the former as the target function of cluster; the latter as $\sum_{c_i, c_j \in C}\lVert c_i - c_j \rVert^2$, in which $C$ is the set of cluster centers.
% }

% \problem{3}{Use hierarchical clustering methods to cluster $Data_i.csv$}
% \solution{Solution}{
% The code of hierarchical clustering and the test script can be found from attachments(hierarchical\_clustering.R, hw1.3.R).
% }
% \solution{Result}{
% It is somehow embarrassing that I still did't know how to draw trees in R without $hcluster()$. The result can be shown is the attachments, (HC\_DATA1.txt and HC\_DATA2.txt). \\[3pt]
% Each result is a (n-1) * narg matrix, in which n is the number of data points and narg is the number of dims in each data point. Each row in the matrix(for example, A[1, ]) means a step when two points merged with each other; A[1, 1] and A[1, 2] merged into a new A[1, 1], and the point represented by A[1, 2] is abandoned. \\[3pt]
% It is possible to draw the tree and do Tree-Cut with the results.
% }

% \refgroup{
% \reference{a a a a  a a}
% \reference{b b b b b b}
% \reference{c c c c c c}
% \reference{b}

% }

\end{document}