### 2016-11-05 Statistical Learning/Homework3/assignment.tex
assignment 3

%!TEX program = xelatex

% \documentclass[10pt]{article}
\input{structure.tex}
\usepackage{epstopdf}
\usepackage{graphics}
\usepackage{subfig}
\usepackage{listings}
\lstset{
  numbers=left,
    framexleftmargin=10mm,
    frame=none,
    backgroundcolor=\color[RGB]{245,245,244},
  keywordstyle=\bf\color{blue},
  identifierstyle=\bf,
  numberstyle=\color[RGB]{0,192,192},
  commentstyle=\it\color[RGB]{0,96,96},
  stringstyle=\rmfamily\slshape\color[RGB]{128,0,0},
  showstringspaces=false,
  extendedchars=false
    }
\DeclareGraphicsExtensions{.eps,.ps,.jpg,.bmp}

\begin{document}

\title{Homework 3}{16.11.4}

% \parbox{0.3\textwidth}{
% Chuan Lu}
% \hfill
% \parbox{0.3\textwidth}{
% 13300180056}
% \hfill
% \parbox{0.3\textwidth}{
% chuanlu13@fudan.edu.cn}

\problem{1}{Perform PCA to assign31.csv.}
\solution{Result}{
Since there are many NAs in assign31.csv, we changed those NA to 0. \n
The results are as follows: The rate in the table is the rate of information reserved after PCA.
\begin{table}[htbp] \large \centering
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
k & 1 & 2 & 3 & 5 & 10 & 14  \\
\hline
Rate & 0.2139058 &  0.396026 & 0.5165856 & 0.6697862 & 0.920202 & 1\\
\hline
\end{tabular}
\caption{Reserved information rate of different orders in PCA.}
\end{table}

\begin{figure}[htbp]
\centering
\includegraphics[width = 15cm]{pics/hw1_pca_1.eps}
\caption{PCA, k = 1}
\label{PCA1}
\end{figure}
\begin{figure}[htbp]
\centering
\includegraphics[width = 15cm]{pics/hw1_pca_2.eps}
\caption{PCA, k = 2}
\label{PCA2}
\end{figure}
\begin{figure}[htbp]
\centering
\includegraphics[width = 15cm]{pics/hw1_pca_3.png}
\caption{PCA, k = 3}
\label{PCA3}
\end{figure}
From these figures(when $k = 1, 2, 3$) we can find that after PCA the data points still seem to be random and messy. When we calculate the rate of reserved information we can know that not until $k = 10$ the rate is less than 0.9, which is thought to be unsatisfying.
}

\problem{2}{Perform PCA and NMF to assign2.csv, and explain the difference.}
\solution{Result}{
For PCA, the result is as follows:
\begin{table}[htbp] \large \centering
\begin{tabular}{|c|c|c|c|}
\hline
k & 1 & 2 & 3  \\
\hline
Rate & 0.4326317 &  0.7648243 & 0.9298461 \\
\hline
\end{tabular}
\caption{Reserved information rate of different orders in PCA.}
\end{table}
\begin{figure}[htbp]
\centering
\includegraphics[width = 15cm]{pics/hw2_pca_1.eps}
\caption{PCA, k = 1}
\label{PCA1}
\end{figure}
\begin{figure}[htbp]
\centering
\includegraphics[width = 15cm]{pics/hw2_pca_2.eps}
\caption{PCA, k = 2}
\label{PCA2}
\end{figure}
\begin{figure}[htbp]
\centering
\includegraphics[width = 15cm]{pics/hw2_pca_3.png}
\caption{PCA, k = 3}
\label{PCA3}
\end{figure}
From figure(k = 1) we can assert nothing; from figure(k = 2) it seems the data can be simulated by two parallel lines. From the table we can find when k = 3, the rate of reserved information is over 0.9, which means k = 3 is a good estimation. \n
For NMF, the result of error rate is as follows:
$$ rate = \frac{\vt{A - WH}_F}{\vt{A}_F} $$
\begin{table}[htbp] \large \centering
\begin{tabular}{|c|c|c|c|c|c|}
\hline
k & 1 & 2 & 3 & 5 & 10\\
\hline
Rate & 0.6764784 &  0.6904373 & 2.629497 & 4.993591 & 14.36354\\
\hline
\end{tabular}
\caption{Error rates of different orders in NMF.}
\end{table}
It must be wrong somewhere that with k being larger, the estimation becoming worse. I think it should be the number of iterations, which is fixed at 1000.
}

\problem{3}{Perform PCA and nmf to test2.csv}
\solution{Result}{
For PCA, the result is shown as follows:
\begin{figure}[htbp]
\centering
\includegraphics[width = 5cm]{pics/hw3_original.png}
\caption{Original Picture}
\label{Original}
\end{figure}

\begin{figure}[htbp]
\centering
\subfloat[k = 1]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw3_pca_1.png}}\hfill
\subfloat[k = 2]{
	\includegraphics[width = 0.3\textwidth]{pics/hw3_pca_2.png}}\hfill
\subfloat[k = 3]{
	\includegraphics[width = 0.3\textwidth]{pics/hw3_pca_3.png}}	
\end{figure}
\begin{figure}[htbp]
\centering
\subfloat[k = 4]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw4_pca_4.png}}\hfill
\subfloat[k = 5]{
	\includegraphics[width = 0.3\textwidth]{pics/hw3_pca_5.png}}\hfill
\subfloat[k = 10]{
	\includegraphics[width = 0.3\textwidth]{pics/hw3_pca_10.png}}	
\end{figure}
\begin{figure}[htbp]
\centering
\subfloat[k = 20]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw3_pca_20.png}}\hfill
\subfloat[k = 50]{
	\includegraphics[width = 0.3\textwidth]{pics/hw3_pca_50.png}}\hfill
\subfloat[k = 200]{
	\includegraphics[width = 0.3\textwidth]{pics/hw3_pca_200.png}}	
\end{figure}
For NMF, 
}
\end{document}