### 2017-03-15 Numerical solution of DEs/Assignment1/Assignment_week_1.tex
update

%!TEX program = xelatex

% \documentclass[10pt]{article}
\input{structure.tex}
\usepackage{epstopdf}
\usepackage{graphics}
\usepackage{subfig}
\usepackage{listings}
\lstset{
  numbers=left,
    framexleftmargin=10mm,
    frame=none,
    backgroundcolor=\color[RGB]{245,245,244},
  keywordstyle=\bf\color{blue},
  identifierstyle=\bf,
  numberstyle=\color[RGB]{0,192,192},
  commentstyle=\it\color[RGB]{0,96,96},
  stringstyle=\rmfamily\slshape\color[RGB]{128,0,0},
  showstringspaces=false,
  extendedchars=false
    }
\DeclareGraphicsExtensions{.eps,.ps,.jpg,.bmp}

\begin{document}

\title{Assignment 1}
\\
Code can be obtained at https://github.com/orcuslc/Learning/

% \parbox{0.3\textwidth}{
% Chuan Lu}
% \hfill
% \parbox{0.3\textwidth}{
% 13300180056}
% \hfill
% \parbox{0.3\textwidth}{
% chuanlu13@fudan.edu.cn}

\problem{1}{Show the numerical result of $f(x) = \frac{1 - cos(x)}{x ^{2}}$ when $x \to 0^+$.}
\solution{Solution}{
The result is as follows.
\begin{figure}[h!]
	\centering
    \includegraphics[width=15cm]{pics/problem1.eps}
    \caption{problem1}
    \label{problem1}
\end{figure}
}

\problem{2}{Analyse the inversed iteration of $u_{n+2} = 3u_{n+1}-2u_n$}
\solution{Solution}{
The inversed iteration is
$$u_n = \frac{3}{2}u_{n+1}-\frac{1}{2}u_{n+2}.$$
And the computed value is
$$\hat{u_n} = \frac{3}{2}\hat{u_{n+1}}-\frac{1}{2}\hat{u_{n+2}}+\epsilon_n.$$
If we let
$$w_n = \left(\begin{array}{cc} u_n \\ u_{n+1}\end{array}\right),A = \left(\begin{array}{cc} \frac{3}{2} & -\frac{1}{2} \\ 1 & 0\end{array}\right), \hat{w_n} = \left(\begin{array}{cc} \hat{u_n} \\ \hat{u_{n+1}}\end{array}\right), \hat{\epsilon_n} = \left(\begin{array}{cc} \epsilon_n \\ 0\end{array}\right)$$
Then
$$w_{n} = Aw_{n+1}, \hat{w_n} = A\hat{w_{n+1}}+\hat{\epsilon_n}$$
Define the error as $e_n = \hat{w_n} - w_n$, then 
$$e_n = Ae_{n+1} + \hat{\epsilon_n}$$
Since $A$ has the eigenvalue decomposition
$$P^{-1}AP = \Lambda, P = \left(\begin{array}{cc} 1 & \frac{1}{2} \\ 1 & 1\end{array}\right), \Lambda = \left(\begin{array}{cc} 1 & 0 \\ 0 & \frac{1}{2}\end{array}\right)$$
the iteration can be wrote as
$$P^{-1}e_n = \Lambda P^{-1}e_{n+1} + P^{-1}\hat{\epsilon_n}$$
With the same analysis of Page 13 of the textbook, we can know that $e_0$ is under control by a constant value of $e_n$.
}

\problem{3}{Prove (1.2.8) of the textbook}
\solution{Proof}{
From (1.2.7) we can know
$$\|x_{n+1} - x_{n}\| \leqslant \alpha^n\|x_1 - x_0\|$$
So
$$\|x^{*} - x_n\| \leqslant \sum_{k = n}^{\infty}\|x_{k+1} - x_k\| \leqslant \sum_{k = n}^{\infty} \alpha^k\|x_1 - x_0\| = \frac{\alpha^n}{1-\alpha}\|x_1 - x_0\|$$
}

\end{document}