### 2017-04-06 Numerical solution of DEs/Assignment4/assignment.tex
add assignment 4&5

%!TEX program = xelatex

% \documentclass[10pt]{article}
\input{structure.tex}
\usepackage{epstopdf}
\usepackage{graphics}
\usepackage{subfig}
\usepackage{listings}
\lstset{
  numbers=left,
    framexleftmargin=10mm,
    frame=none,
    backgroundcolor=\color[RGB]{245,245,244},
  keywordstyle=\bf\color{blue},
  identifierstyle=\bf,
  numberstyle=\color[RGB]{0,192,192},
  commentstyle=\it\color[RGB]{0,96,96},
  stringstyle=\rmfamily\slshape\color[RGB]{128,0,0},
  showstringspaces=false,
  extendedchars=false
    }
\DeclareGraphicsExtensions{.eps,.ps,.jpg,.bmp}

\begin{document}

\title{Assignment 4}{17.4.1}

\problem{1}{Prove Implicit Euler Iteration is of first-order convergence for Dahlquist test.}
\solution{Proof}{
\quad According to (2.2.20), $u_{n} = \frac{1}{(1-a\Delta t)^{n}}u_{0}.$ For $e_{n} = u(t_{n}) - u_{n},$ we have $e_{n} = u_{0}(e^{at_{n}} - e^{-nln(1-a\Delta t)}).$ From Taylor expansion we can know that $$e_{n} = -u_{0}e^{at_{n}}(\frac{a^{2}t_{n}}{2}\Delta t + O(\Delta t^{2})).$$\\
\quad In consequence, for a fixed $T = N\Delta t,$ $\Vert u(T) - u(N)\Vert\leqslant C\Delta t,$ which means the implicit iteration is of first-order convergence.
}

\problem{2}{Compute $\frac{du}{dt} = au$ with Euler Iterations, and show the convergence by image.}
\solution{Result}{
\begin{figure}[!h]
\centering
\includegraphics[width = 15cm]{img/converge.eps}
\caption{The convergence of Euler Iteration.}
\label{Convergence}
\end{figure}
\begin{figure}[!h]
\centering
\includegraphics[width = 15cm]{img/error_list.eps}
\caption{The error of Euler Iteration.}
\label{Error}
\end{figure}
}

\end{document}