### 2017-04-29 Numerical solution of DEs/Assignment7/assignment.tex
Ass Assignment 7

%!TEX program = xelatex

% \documentclass[10pt]{article}
\input{structure.tex}
\usepackage{epstopdf}
\usepackage{graphics}
\usepackage{subfig}
\usepackage{listings}
\lstset{
  numbers=left,
    framexleftmargin=10mm,
    frame=none,
    backgroundcolor=\color[RGB]{245,245,244},
  keywordstyle=\bf\color{blue},
  identifierstyle=\bf,
  numberstyle=\color[RGB]{0,192,192},
  commentstyle=\it\color[RGB]{0,96,96},
  stringstyle=\rmfamily\slshape\color[RGB]{128,0,0},
  showstringspaces=false,
  extendedchars=false
    }
\DeclareGraphicsExtensions{.eps,.ps,.jpg,.bmp}

\begin{document}

\title{Assignment 7}{17.4.26}

\problem{1}{Derive A-B, A-M and Gear Formula with Newton Interpolation.}
\solution{Solution}{
With Newton Interpolation on $t_{n+1}, t_{n}, \cdots, t_{n+1-k}$, $f(t, u) = f_{n+1-k} + f_{n+1-k, n+2-k}(t - t_{n+1-k}) + \dots + f_{n+1-k, n+2-k ,\cdots, n+1}\Pi_{i=1}^{k}(t-t_{n-k+i}) + f_{n+1-k, n+2-k, \cdots, n+1, t}\Pi_{i=1}^{k+1}(t-t_{n-k+i}).$ \\
Let $p_{n, k}(t) = f(t, u) - f_{n+1-k, n+2-k, \cdots, n+1, t}\Pi_{i=1}^{k+1}(t-t_{n-k+i})$, 
Integrate within $[t_{n}, t_{n+1}]$, $u_{n+1} - u_{n} = \Delta t\Sigma_{i=0}^{k}b_{k, i}f_{i}$. 
\subsection{Adams-Bashforth Formula}
Like Adams-Moulton formula, with Newton interpolation on $t_{n}, t_{n-1}, \cdots, t_{n-k}$, $f(t, u) = f_{n-k} + f_{n-k, n+1-k}(t - t_{n-k}) + \dots + f_{n-k, n+1-k ,\cdots, n}\Pi_{i=0}^{k-1}(t-t_{n-k+i}) + f_{n-k, n+1-k, \cdots, n, t}\Pi_{i=0}^{k}(t-t_{n-k+i}).$ \\
Let $p_{n, k}(t) = f(t, u) - f_{n-k, n+1-k, \cdots, n, t}\Pi_{i=0}^{k}(t-t_{n-k+i})$, 
Integrate within $[t_{n}, t_{n+1}]$, $u_{n+1} - u_{n} = \Delta t\Sigma_{i=0}^{k}b_{k, i}f_{i}$. 
\subsection{Gear Formula}
Let $I_{k}u = u_{n-k+1} + u_{n-k+1, n-k+2}(t - t_{n-k+1}) + \cdots + u_{n-k+1, n-k+2, \cdots, n+1}\Pi_{i = n-k+1}^{n}(t-t_{i})$. \\
Use derivative of $I_{n}u(t)$ to take the place of the derivative of $u(t)$ when $t = t_{n+1}$, there exists $$f(t, u)|_{t = t_{n+1}} = \frac{du}{dt}|_{t=t_{n+1}} = \frac{dI_{k}u}{dt}|_{t=t_{n+1}}$$. \\
In consequence, $\Delta tf_{n+1} = (\Delta t)^{2}(u_{n-k+1} + u_{n-k+1, n-k+2}(t - t_{n-k+1}) + \cdots + u_{n-k+1, n-k+2, \cdots, n+1}\Pi_{i = n-k+1}^{n}(t-t_{i}))$.
}

\problem{2}{Consider the equation $$\frac{du}{dt}=\lambda(-u + cos(t))$$}
\solution{Solution}{

\subsection{The exact solution of the equation.}
The exact solution is $u(t) = c_{0}e^{-\lambda t}+\int_{0}^{t}e^{-\lambda(t-s)}\lambda cos(s)ds$.

\subsection{For $\lambda = 1,10,100,1000,$ use explicit and implicit Euler iteration to solve.}
\begin{figure}[!h]
\centering
\includegraphics[width = 15cm]{lambda=1.eps}
\caption{lambda = 1}
\label{Lambda1}
\end{figure}
\begin{figure}[!h]
\centering
\includegraphics[width = 15cm]{lambda=10.eps}
\caption{lambda = 10}
\label{Lambda2}
\end{figure}
\begin{figure}[!h]
\centering
\includegraphics[width = 15cm]{lambda=100.eps}
\caption{lambda = 100}
\label{Lambda3}
\end{figure}
\begin{figure}[!h]
\centering
\includegraphics[width = 15cm]{lambda=1000.eps}
\caption{lambda = 1000}
\label{Lambda4}
\end{figure}

\subsection{Use Adams and Gear iteration to solve the equation when $\lambda=1000$.}
\begin{figure}[!h]
\centering
\includegraphics[width = 15cm]{3.eps}
\caption{Adams and Gear}
\label{Lambda}
\end{figure}
.
}

\end{document}