### 2017-09-22 Graduate - Numerical Analysis/Assignment 5/assignment.tex
update

%!TEX program = xelate
\input{structure.tex}
\usepackage{epstopdf}
\usepackage{graphics}
\usepackage{subfig}
\usepackage{listings}
\lstset{
  breaklines=true,
  xleftmargin=25pt,
  xrightmargin=25pt,
  aboveskip=0pt,
  belowskip=10pt,
  basicstyle=\ttfamily,
  showstringspaces=false,
  frame=ltrb,
  tabsize=4,
  numbers=left,
  numberstyle=\small,
  numbersep=8pt,
  morekeywords={*, factorial, sum, erlang},
}
\DeclareGraphicsExtensions{.eps,.ps,.jpg,.bmp}

\begin{document}

\title{Numerical Analysis \\ Assignment 5}
\date{\today}
\author{Chuan Lu}

\maketitle

\problem{1}{Problem 2.21, Page 121}
\solution{Solution}{
According to Theorem 2.7, we should choose $c$ so as to have $|g'(\alpha)| < 1$. Thus
$$
|g'(\alpha)| = |1+cf'(\alpha)| < 1.
$$
So $c$ satisfies $-2 < cf'(\alpha) < 0$. For a good rate of convergence, pick $c$ s.t. $cf'(\alpha) \sim 0$.
}

\problem{2}{Problem 2.24, Page 121}
\solution{Solution}{
\textbf{(a)} 
$$
g(x) = -16+6x+\frac{12}{x}.
$$
Thus $g'(2) = 3 > 1.$ So the iteration does not converge.

\textbf{(b)}
$$
g(x) = \frac{2}{3}x+\frac{1}{x^2}.
$$
Thus $g'(3^{\frac{1}{3}}) = \frac{2}{3}-2\alpha^{-3} = 0$. So the iteration converges. Since
$$
\begin{aligned}
\lim_{n\to\infty}\frac{\alpha-x_{n+1}}{(\alpha-x_n)^2} &= \lim_{n\to\infty}\frac{-2x_n^3+3\alpha x_n^2-3}{3x_n^2(\alpha-x_n)^2} = \lim_{x_n\to\alpha}\frac{-2x_n^3+3\alpha x_n^2-3}{3x_n^2(\alpha-x_n)^2}\\
&= \lim_{x_n\to\alpha}\frac{-6x_n^2+6\alpha x_n}{6x_n(\alpha-x_n)^2-6x_n^2(\alpha-x_n)} = \lim_{x_n\to\alpha}\frac{1}{\alpha-2x_n} \\
&= -\frac{1}{\alpha}.
\end{aligned}
$$
We know the iteration is second-order convergent.

\textbf{(c)}
$$
g(x) = \frac{12}{1+x}.
$$
Thus $g'(\alpha) = -\frac{3}{4}$, so the iteration converges. Since
$$
\lim_{n\to\infty}\frac{\alpha-x_{n+1}}{\alpha-x_n} = \frac{(1+x_n)\alpha-12}{(1+x_n)(\alpha-x_n)} = \frac{\alpha}{\alpha-2x_n-1} = -\frac{3}{4},
$$
We know the iteration is linear convergence, and the rate is $\frac{3}{4}$.
}

\problem{3}{Problem 2.28, Page 122}
\solution{Proof}{
Assume $f(x) = (x-\alpha)h(x), ~h(\alpha) \ne 0$. Then
$$
\begin{aligned}
g(x) &= x - \frac{f(x)}{D(x)} = x-\frac{(x-\alpha)^2h^2(x)}{(x-\alpha)(h(x)+1)h((x-\alpha)h(x)+x)-(x-\alpha)h(x)} \\
&= x - \frac{(x-\alpha)h^2(x)}{(h(x)+1)h((x-\alpha)h(x)+x)-h(x)}.
\end{aligned}
$$
Thus 
$$
\begin{aligned}
g'(\alpha) &= \left.1-\frac{(h^2(x)+2(x-\alpha)h(x)h'(x))((h(x)+1)h((x-\alpha)h(x)+x)-h(x)) - (x-\alpha)M}{((h(x)+1)h((x-\alpha)h(x)+x)-h(x))^2}\right|_{x=\alpha} \\
&= 1-\frac{h^4(\alpha)}{h^4(\alpha)} = 0.
\end{aligned}
$$
And we can know that $g''(\alpha)\ne 0$. Accoding to Theorem 2.8, this iteration is a second-order method.
}

\problem{4}{Problem 2.48, Page 126}
\solution{Proof}{
According to the continuity of $\lVert\cdot\rVert_{\infty}$, we can find an $\epsilon$ and a closed set $B(\alpha, \delta)$, s.t. $\forall x\in B(\alpha, \delta)$, $\lVert G(x)\rVert_{\infty} \le 1-\epsilon$. Then $\forall x_0 \in B = B(\alpha, \delta)$,
$$
\lVert \alpha-x_{n+1}\rVert = \lVert g(\alpha)-g(x_n)\rVert = \lVert G(\xi)\rVert \cdot \lVert \alpha-x_n\rVert \le (1-\epsilon) \lVert \alpha-x_n\rVert \le (1-\epsilon)^n \lVert \alpha-x_0\rVert.
$$
Thus $\lVert \alpha-x_{n+1}\rVert \to 0, ~t\to\infty$. So the iteration converges to $\alpha$. And since $g(B) \subset B$ according to that $\lVert\alpha-g(x)\rVert < \lVert\alpha-x\rVert$, and $\max \lVert G(x)\rVert < 1$, we know that it satisfies the condition of Theorem 2.9.
}

\problem{5}{Problem 2.50, Page 127}
\solution{Solution}{

}
\end{document}