### 2017-09-23 Graduate - Numerical Analysis/Assignment 5/assignment.tex
Add HW 5

%!TEX program = xelate
\input{structure.tex}
\usepackage{epstopdf}
\usepackage{graphics}
\usepackage{subfig}
\usepackage{listings}
\lstset{
  breaklines=true,
  xleftmargin=25pt,
  xrightmargin=25pt,
  aboveskip=0pt,
  belowskip=10pt,
  basicstyle=\ttfamily,
  showstringspaces=false,
  frame=ltrb,
  tabsize=4,
  numbers=left,
  numberstyle=\small,
  numbersep=8pt,
  morekeywords={*, factorial, sum, erlang},
  keywordstyle=\color{blue!70}, commentstyle=\color{red!50!green!50!blue!50},
}
\DeclareGraphicsExtensions{.eps,.ps,.jpg,.bmp}

\begin{document}

\title{Numerical Analysis \\ Assignment 5}
\date{\today}
\author{Chuan Lu}

\maketitle

\problem{1}{Problem 2.21, Page 121}
\solution{Solution}{
According to Theorem 2.7, we should choose $c$ so as to have $|g'(\alpha)| < 1$. Thus
$$
|g'(\alpha)| = |1+cf'(\alpha)| < 1.
$$
So $c$ satisfies $-2 < cf'(\alpha) < 0$. For a good rate of convergence, pick $c$ s.t. $cf'(\alpha) \sim 0$.
}

\problem{2}{Problem 2.24, Page 121}
\solution{Solution}{
\textbf{(a)} 
$$
g(x) = -16+6x+\frac{12}{x}.
$$
Thus $g'(2) = 3 > 1.$ So the iteration does not converge.

\textbf{(b)}
$$
g(x) = \frac{2}{3}x+\frac{1}{x^2}.
$$
Thus $g'(3^{\frac{1}{3}}) = \frac{2}{3}-2\alpha^{-3} = 0$. So the iteration converges. Since
$$
\begin{aligned}
\lim_{n\to\infty}\frac{\alpha-x_{n+1}}{(\alpha-x_n)^2} &= \lim_{n\to\infty}\frac{-2x_n^3+3\alpha x_n^2-3}{3x_n^2(\alpha-x_n)^2} = \lim_{x_n\to\alpha}\frac{-2x_n^3+3\alpha x_n^2-3}{3x_n^2(\alpha-x_n)^2}\\
&= \lim_{x_n\to\alpha}\frac{-6x_n^2+6\alpha x_n}{6x_n(\alpha-x_n)^2-6x_n^2(\alpha-x_n)} = \lim_{x_n\to\alpha}\frac{1}{\alpha-2x_n} \\
&= -\frac{1}{\alpha}.
\end{aligned}
$$
We know the iteration is second-order convergent.

\textbf{(c)}
$$
g(x) = \frac{12}{1+x}.
$$
Thus $g'(\alpha) = -\frac{3}{4}$, so the iteration converges. Since
$$
\lim_{n\to\infty}\frac{\alpha-x_{n+1}}{\alpha-x_n} = \frac{(1+x_n)\alpha-12}{(1+x_n)(\alpha-x_n)} = \frac{\alpha}{\alpha-2x_n-1} = -\frac{3}{4},
$$
We know the iteration is linear convergence, and the rate is $\frac{3}{4}$.
}

\problem{3}{Problem 2.28, Page 122}
\solution{Proof}{
Assume $f(x) = (x-\alpha)h(x), ~h(\alpha) \ne 0$. Then
$$
\begin{aligned}
g(x) &= x - \frac{f(x)}{D(x)} = x-\frac{(x-\alpha)^2h^2(x)}{(x-\alpha)(h(x)+1)h((x-\alpha)h(x)+x)-(x-\alpha)h(x)} \\
&= x - \frac{(x-\alpha)h^2(x)}{(h(x)+1)h((x-\alpha)h(x)+x)-h(x)}.
\end{aligned}
$$
Thus 
$$
\begin{aligned}
g'(\alpha) &= \left.1-\frac{(h^2(x)+2(x-\alpha)h(x)h'(x))((h(x)+1)h((x-\alpha)h(x)+x)-h(x)) - (x-\alpha)M}{((h(x)+1)h((x-\alpha)h(x)+x)-h(x))^2}\right|_{x=\alpha} \\
&= 1-\frac{h^4(\alpha)}{h^4(\alpha)} = 0.
\end{aligned}
$$
And we can know that $g''(\alpha)\ne 0$. Accoding to Theorem 2.8, this iteration is a second-order method.
}

\problem{4}{Problem 2.48, Page 126}
\solution{Proof}{
According to the continuity of $\lVert\cdot\rVert_{\infty}$, we can find an $\epsilon$ and a closed set $B(\alpha, \delta)$, s.t. $\forall x\in B(\alpha, \delta)$, $\lVert G(x)\rVert_{\infty} \le 1-\epsilon$. Then $\forall x_0 \in B = B(\alpha, \delta)$,
$$
\lVert \alpha-x_{n+1}\rVert = \lVert g(\alpha)-g(x_n)\rVert = \lVert G(\xi)\rVert \cdot \lVert \alpha-x_n\rVert \le (1-\epsilon) \lVert \alpha-x_n\rVert \le (1-\epsilon)^n \lVert \alpha-x_0\rVert.
$$
Thus $\lVert \alpha-x_{n+1}\rVert \to 0, ~t\to\infty$. So the iteration converges to $\alpha$. And since $g(B) \subset B$ according to that $\lVert\alpha-g(x)\rVert < \lVert\alpha-x\rVert$, and $\max \lVert G(x)\rVert < 1$, we know that it satisfies the condition of Theorem 2.9.
}

\problem{5}{Problem 2.50, Page 127}
\solution{Solution}{
The code and result are as follows;(In fact)
}
\lstinputlisting[language = MATLAB]{Newton.m}
\lstinputlisting[language = MATLAB]{prob50.m}
\begin{lstlisting}[language = MATLAB]
>> prob50
   1.600000000000000   1.200000000000000
   1.581250000000000   1.225000000000000
   1.581138833992095   1.224744897959184
\end{lstlisting}

\problem{6}{Problem 2.51, Page 127}
\solution{Solution}{
We just use the code of Newton's method from Problem 2.50, and the executive script is as follows;
}
\lstinputlisting[language = MATLAB]{prob51.m}
\solution{Result}{
The result is shown in the following table \ref{Table 1}, and the convergence speed shown in \ref{Graph 1}. From the result we know that different initial values may lead to convergence to different roots, while when the iteration comes to very close to the root, the convergence order is first-order.
}
\begin{table}[!htb]
\centering
\begin{tabular}{|c|c|c|c|}
\hline
Initial Point & Convergence & Result & Number of iterates \\
\hline
(1.2, 2.5) & Yes & (1.336355248159268, 1.754234729131226) & 11 \\
\hline
(-2, 2.5) & Yes & (1.336355247361990, 1.754234726236878) & 19 \\
\hline
(-1.2, -2.5) & Yes & (1.336355235838950, 1.754234684404958) & 24 \\
\hline
(2, -2.5) & Yes & (2.998365352785336, 0.148430977265564) & 9 \\
\hline
\end{tabular}
\caption{The result of convergence and iteration count.}
\label{Table 1}
\end{table}

\begin{figure}[!htb]
\centering
\begin{minipage}{0.45\textwidth}
\includegraphics[width = \textwidth]{res1.eps}
\end{minipage}
\begin{minipage}{0.45\textwidth}
\includegraphics[width = \textwidth]{res2.eps}
\end{minipage} \\
\begin{minipage}{0.45\textwidth}
\includegraphics[width = \textwidth]{res3.eps}
\end{minipage}
\begin{minipage}{0.45\textwidth}
\includegraphics[width = \textwidth]{res4.eps}
\end{minipage}
\caption{Convergence Order}
\label{Figure 1}
\end{figure}
\end{document}