### 2017-10-09 Graduate - Introduction to Analysis I/Assignment 5/5-Homework20171008.tex
add prob 1

\documentclass{article}%
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amsfonts}%
\usepackage{amssymb}

\setlength{\topmargin}{-0.75in}
\setlength{\textheight}{9.25in}
\setlength{\oddsidemargin}{0.0in}
\setlength{\evensidemargin}{0.0in}
\setlength{\textwidth}{6.5in}
\def\labelenumi{\arabic{enumi}.}
\def\theenumi{\arabic{enumi}}
\def\labelenumii{(\alph{enumii})}
\def\theenumii{\alph{enumii}}
\def\p@enumii{\theenumi.}
\def\labelenumiii{\arabic{enumiii}.}
\def\theenumiii{\arabic{enumiii}}
\def\p@enumiii{(\theenumi)(\theenumii)}
\def\labelenumiv{\arabic{enumiv}.}
\def\theenumiv{\arabic{enumiv}}
\def\p@enumiv{\p@enumiii.\theenumiii}
\pagestyle{plain}
\setcounter{secnumdepth}{0}
\newtheorem{theorem}{Theorem}
\newtheorem{acknowledgement}[theorem]{Acknowledgement}
\newtheorem{algorithm}[theorem]{Algorithm}
\newtheorem{axiom}[theorem]{Axiom}
\newtheorem{case}[theorem]{Case}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{conclusion}[theorem]{Conclusion}
\newtheorem{condition}[theorem]{Condition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{criterion}[theorem]{Criterion}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{solution}[theorem]{Solution}
\newtheorem{summary}[theorem]{Summary}
\newenvironment{proof}[1][Proof]{\textbf{#1.} }{\ \rule{0.5em}{0.5em}}

\begin{document}

\begin{center}
\textbf{Introduction to Analysis $I$\\Homework 5\\Sunday, October 8, 2017}\bigskip
\end{center}

\noindent\textbf{Instructions}: This and all subsequent homeworks must be submitted written in \LaTeX.

\noindent If you use results from books, Royden or others, please be explicit about what results you are using.

\begin{center}
\emph{Homework 5 is due by midnight, Saturday, October 21.}
\end{center} 
\medskip

\begin{enumerate}
\item  (Problem 24, Page 64) Let $I$ be an interval in $\mathbb{R}$ and let $f:I\to \mathbb{R}$ be increasing. Show that $f$ is measurable by first showing that, for each natural number $n$, the strictly increasing function $x\to f(x)+x/n$ is measurable, and then taking pointwise limits.

\bigskip
\textbf{Collaborators:} None\\
\smallskip

\textbf{Solution:}
Denote $f_{n}(x) = f(x)+\frac{x}{n}$. Then since $f: I \to \mathbb{R}$ is increasing, $f_n(x)$ is strictly increasing on $I \in \{[a, b], [a, b), (a, b], (a, b)\}$.

For each fixed number $c$, if $\exists x_0\in I$, s.t. $f(x_0) = c$, then the set
$$
\{x\mid f(x) < c\} = (a, x_0),
$$
and the left side is the same as the set $I$, and it is an interval in $\mathbb{R}$, hence is measurable. Thus $f_{n}$ is measurable for every $n$.

For each $x\in I$, 
$$
\lim_{n\to\infty}f_n(x) = f(x)+\lim_{n\to\infty}\frac{x}{n} = f(x),
$$
hence $f_{n}$ converges to $f$ pointwise. Using Proposition 9, we know $f$ is measurable.
\bigskip

\item (Problem 8, Page 343)  Let $(X,\mathcal{M}, \mu)$ be a measure space.  The measure $\mu$ is said to be \textbf{semifinite} provided each measurable set of infinite measure contains measurable sets of arbitrarily large finite measure.
\begin{enumerate}
\item Show that each $\sigma$-finite measure is semifinite.
\item For $E\in \mathcal{M}$, define $\mu_1(E) = \mu(E)$, if $\mu(E)<\infty$, and if $\mu(E)=\infty$, define $\mu_1(E)=\infty$ if $E$ contains measurable sets of arbitrarily large finite measure and $\mu_1(E)= 0$ otherwise.  Show $\mu_1$ is a semifinite measure: it is called the semifinite part of $\mu$.
\item Find a measure $\mu_2$ on $\mathcal{M}$ that only takes the valures $0$ and $\infty$ and $\mu = \mu_1 + \mu_2$.
\end{enumerate}

\bigskip
\textbf{Collaborators:}\\
\smallskip

\textbf{Solution:}
\bigskip

\item (Problem 9, Page 343) Prove Proposition 3, that is, show that $\mathcal{M}_0$ is a $\sigma$-algebra, $\mu_0$ is properly defined and $(X,\mathcal{M}_0, \mu_0)$ is complete.  In what sense is $\mathcal{M}_0$ minimal?

\bigskip
\textbf{Collaborators:}\\
\smallskip

\textbf{Solution:}
\bigskip

\item (Problem 10, Page 343) If $(X,\mathcal{M},\mu)$ is a measure space, we say that a subset $E$ of $X$ is \textbf{locally measurable} provided for each $B\in \mathcal{M}$ with $\mu(B)<\infty$, the intersection $E\cap B$ belongs to $\mathcal{M}$.  The measure $\mu$ is called \textbf{saturated} provided every locally measurable set is measurable.
\begin{enumerate}
\item Show that each $\sigma$-finite measure is saturated.
\item Show that the collection $\mathcal{C}$ of locally measurable sets is a $\sigma$-algebra.
\item Let $(X,\mathcal{M},\mu)$ be a measure space and $\mathcal{C}$ the $\sigma$ of locally measurable sets. For $E\in \mathcal{C}$, define $\overline{\mu}(E) = \mu(E)$ if $E\in \mathcal{M}$ and $\overline{\mu}(E) = \infty$ if $E\notin \mathcal{M}$.  Show that $(X,\mathcal{C},\overline{\mu})$ is a saturated measure space.
\item If $\mu$ is semifinite and $E\in \mathcal{C}$, set $\underline{\mu}(E)=\sup\{\mu(B)\mid B\in\mathcal{M},\, B\subseteq E \}$. Show that  $(X,\mathcal{C},\underline{\mu})$ is a saturated measure space and that $\underline{\mu}$ is an extension of $\mu$.  Give an example to show that $\overline{\mu}$ and $\underline{\mu}$ may be different.
\end{enumerate}

\bigskip
\textbf{Collaborators:}\\
\smallskip

\textbf{Solution:}
\bigskip

\item (Problem 18, Page 373)  Let $\{u_n\}$ be a sequence of nonnegative measurable functions on $X$.  For $x\in X$, define $f(x) = \sum_{n=1}^{\infty} u_n(x)$.  Show that \[
\int_X f \, d\mu = \sum_{n=1}^{\infty} \left[\int_X u_n \, d\mu \right].\]

\end{enumerate}
\end{document}