### 2017-10-28 Graduate - Numerical Analysis/Assignment 10/assignment.tex
add prob 3, 4

%!TEX program = xelate
\input{structure.tex}
\usepackage{epstopdf}
\usepackage{graphics}
\usepackage{subfig}
\usepackage{listings}
\lstset{
  breaklines=true,
  xleftmargin=25pt,
  xrightmargin=25pt,
  aboveskip=0pt,
  belowskip=10pt,
  basicstyle=\ttfamily,
  showstringspaces=false,
  frame=ltrb,
  tabsize=4,
  numbers=left,
  numberstyle=\small,
  numbersep=8pt,
  morekeywords={*, factorial, sum, erlang},
  keywordstyle=\color{blue!70}, commentstyle=\color{red!50!green!50!blue!50},
}
\DeclareGraphicsExtensions{.eps,.ps,.jpg,.bmp}

\begin{document}

\title{Numerical Analysis \\ Assignment 9}
\date{\today}
\author{Chuan Lu}

\maketitle

\problem{1}{Problem 4.16, Page 242}
\solution{Solution}{
For each $m < n$, by integration by parts, 
$$
\begin{aligned}
\int_{0}^{\infty}e^{-x} x^m \varphi_n(x)dx &= \frac{(-1)^n}{n!}\int_{0}^{\infty} x^m\frac{d^n}{dx^n}(x^ne^{-x})dx \\
&= \frac{(-1)^n}{n!}\left.\left(x^m\frac{d^{n-1}}{dx^{n-1}}(x^ne^{-x})\right|_{0}^{\infty}-\int_{0}^{\infty}mx^{m-1}\frac{d^{n-1}}{dx^{n-1}}(x^ne^{-x})dx \right)
\end{aligned}
$$
Since 
$$
x^m\frac{d^{n-1}}{dx^{n-1}}(x^ne^{-x}) = e^{-x}N(x),
$$
where $N(x)$ is a polynomial of degree $n-1+m$, by L'Hospital's Rule we know the first term in the integration is $0$. Then by induction we know
$$
\begin{aligned}
\int_{0}^{\infty}e^{-x} x^m \varphi_n(x)dx &= \frac{(-1)^{n+1}m}{n!}\int_{0}^{\infty}x^{m-1}\frac{d^{n-1}}{dx^{n-1}}(x^n e^{-x})dx \\ 
&= \frac{(-1)^{n+m}m!}{n!}\int_{0}^{\infty}\frac{d^{n-m}}{dx^{n-m}}(x^n e^{-x})dx \\
&= \frac{(-1)^{n+m}m!}{n!}\frac{d^{n-m}}{dx^{n-m}}\int_{0}^{\infty}x^n e^{-x}dx \\
&= \frac{(-1)^{n+m}m!}{n!}\frac{d^{n-m}}{dx^{n-m}} (n!) = 0.
\end{aligned}
$$
In the deduction we used the property that $f(x) = x^n e^{-x}$ is absolutely continuous. Since $\varphi_m(x)$ is a polynomial of degree $m < n$, we know 
$$
(\varphi_n(x), \varphi_m(x)) = 0, ~\text{and}~ (\varphi_n(x), \varphi_n(x)) \ne 0.
$$
Hence $\{\varphi_n(x)\}$ is a family of orthogonal polynomials.
}

\problem{2}{Problem 4.18, Page 242}
\solution{Solution}{
First, we derive $c_n$. Multiply both sides of (4.4.21) by $w(x)\varphi_{n-1}(x)$, and then integrate, we get
$$
\int w\varphi_{n+1}\varphi_{n-1}dx = \int a_n wx\varphi_n\varphi_{n-1} + \int b_nw\varphi_n\varphi_{n-1} - c_n\int w\varphi_{n-1}^2.
$$
Using the orthogonality of $\varphi_n$, the left side is 0, and the second term of right side is 0. Then 
$$
a_n\int wx\varphi_n\varphi_{n-1} = c_n\int w\varphi_{n-1}^2.
$$
Since
$$
a_n\int wx\varphi_n\varphi_{n-1} = a_n\int  w\varphi_n(A_{n-1}x^n+B_{n-1}x^{n-1}+\cdots) = a_n\int w\varphi_n A_{n-1}x^n = a_n \frac{A_{n-1}}{A_n}\int w\varphi_n^2,
$$
we have
$$
c_n = \frac{a_n A_{n-1}\gamma_n}{A_n\gamma_{n-1}} = \frac{A_{n+1}A_{n-1}\gamma_n}{A_n^2\gamma_{n-1}}.
$$
Now we consider $b_n$. Multiply both sides of (4.4.21) by $w(x)\varphi_n(x)$, then integrate both sides, we get
$$
\int w\varphi_{n+1}\varphi_n = \int a_n wx\varphi_n^2+ \int b_n w\varphi_n^2 - \int c_n w\varphi_{n-1}\varphi_n.
$$
Using orthogonality, we get
$$
\int a_n wx\varphi_n^2 + \int b_n w\varphi_n^2 = 0.
$$
The first term can be wrote as 
$$
\begin{aligned}
\int a_n wx\varphi_n^2 &= a_n\int w(A_n x^{n+1}+B_n x+\cdots)\varphi_n = a_n \int w (\frac{A_n}{A_{n+1}}\varphi_{n+1}- \frac{A_n B_{n+1}-A_{n+1}B_n}{A_{n+1}}x^n+\cdots)\varphi_n \\
&= a_n\int w(B_n -\frac{A_n}{A_{n+1}}B_{n+1})x^n\varphi_n = a_n\int w\frac{1}{A_n}(B_n -\frac{A_n}{A_{n+1}}B_{n+1})\varphi_n^2.
\end{aligned}
$$
Thus
$$
a_n (\frac{B_n}{A_n}-\frac{B_{n+1}}{A_{n+1}}) \gamma_n + b_n\gamma_n = 0,
$$
we know
$$
b_n = a_n (\frac{B_{n+1}}{A_{n+1}} - \frac{B_n}{A_n}).
$$
}

\problem{3}{Problem 4.21, Page 243}
\solution{Proof}{
Denote $\varphi_n(x) = A_n x^n+B_n x^{n-1}+\cdots $, and $A_n > 0$. Then by (4.4.21), 
$$
\varphi_{n+1}(x) = (a_n x+b_n)\varphi_n(x)-c_n\varphi_{n-1}(x).
$$
We add a $\varphi_0(x)$ to this series, and $\varphi_{0}(x) = A_0 > 0$.
First, when $n = 1$, since
$$
\int_{a}^{b} w(x)\varphi_1(x)\varphi_0(x)dx = \int_{a}^{b}A_0w(x)\varphi_1(x)dx = 0 = A_0\varphi_1(\xi)\int_{a}^{b}w(x) dx,
$$
we know $\varphi_1(\xi) = 0, ~\xi\in (a, b)$. Then we show $\varphi_2(x)$ has two different roots in $(a, b)$. First,
$$
\varphi_2(\xi) = (a_1\xi+b_1)\varphi_1(\xi)-c_1\varphi_0(\xi) = -c_1\varphi_0(\xi) < 0.
$$
Suppose $\varphi_2(x)$ does not change sign in $(a, b)$, then
$$
\int w(x)\varphi_2(x)\varphi_0(x) = A_0\int w(x) \varphi_2(x) < 0.
$$
It makes a contradiction with orthogonality. Then there $\exists x_1\in(a, b) $, s.t. $\varphi_2(x_1) = 0$. Since $A_2 > 0$, $x_1$ cannot be a double root. If $\varphi_2(x)$ has only one root in $(a, b)$, then 
$$
\varphi_2(x)(x-x_1) = q(x)(x-x_1)^2,
$$
integrate by $w(x)$, then since $(x-x_1)$ is of degree 1, left side is 0. But we know $q(x)$ has no root in $(a, b)$, it does not change sign in $(a, b)$, thus the integration is not 0. It makes a contradiction. Thus $\varphi_2(x)$ has two different roots in $(a, b)$, and with $\varphi_2(\xi) < 0$ and $A_2 > 0$, we know the two roots are in $(a, \xi)$ and $(\xi, b)$ seperately. Using the same method, we know that $\varphi_n(x)$ has $n$ different roots in $(a, b)$.

Now we assume this proposition holds for $\varphi_m(x)$, $m \le n$. Denote roots of $\varphi_{n}(x)$ to be $x_i$, then since
$$
\varphi_{n+1}(x_i) = (a_n x_i+b_n)\varphi_n(x_i)-c_n\varphi_{n-1}(x_i) = -c_n\varphi_{n-1}(x_i),
$$
$$
\varphi_{n+1}(x_{i+1}) = (a_n x_{i+1}+b_n)\varphi_n(x_{i+1})-c_n\varphi_{n-1}(x_{i+1}) = -c_n\varphi_{n-1}(x_{i+1}),
$$
from the assumption we know $\varphi_{n-1}(x_i) $ and $\varphi_{n-1}(x_{i+1}) $ has different signs, which means $\varphi_n(x) $ has a root in each of this intervals. Then from $\varphi_n $ has $n$ different roots in $(a, b)$, from induction we get the result.
}

\problem{4}{Problem 4.23, Page 243}
\solution{Solution}{
We know the orthogonal functions on $[-1, 1]$ with weight function $w(x) = \frac{1}{\sqrt{1-x^2}}$ are Chebyshev polynomials:
$$
T_0(x) = 1, ~T_1(x) = x, ~T_2(x) = 2x^2-1.
$$
For any $p(x) = a_0+a_1x+a_2x^2 $ of degree 2 that minimizes the distance, we have
$$
a_j = (f, T_j), ~j = 0, 1, 2.
$$
Thus
$$
a_0 = (f, T_0) = \int_{-1}^{1} \frac{1}{\sqrt{1-x^2}}\cos^{-1}xdx = \int_{\pi}^{0}-ydy = \frac{\pi^2}{2},
$$
$$
a_1 = (f, T_1) = \int_{-1}^{1}\frac{x}{\sqrt{1-x^2}}\cos^{-1}xdx = \left.\int_{0}^{\pi} y\cos ydy = y\sin y\right|_{0}^{\pi} - \int_{0}^{\pi} \sin ydy = -2,
$$
$$
a_2 = (f, T_2) = \int_{-1}^{1}\frac{2x^2-1}{\sqrt{1-x^2}}\cos^{-1}xdx = \left.\int_{0}^{\pi}y\cos 2ydy = \frac{y\sin 2y}{2}\right|_{0}^{\pi} - \frac{1}{2}\int_{0}^{\pi}\sin 2ydy = -\frac{1}{2}.
$$
Thus
$$
p_2(x) = -\frac{1}{2}x^2-2x+\frac{\pi^2}{2}.
$$
}
\end{document}