### 2017-10-29 Graduate - Introduction to Analysis I/Assignment 6/assignment.tex
update

%!TEX program = xelate
\input{structure.tex}
\usepackage{epstopdf}
\usepackage{graphics}
\usepackage{subfig}
\usepackage{listings}
\lstset{
  breaklines=true,
  xleftmargin=25pt,
  xrightmargin=25pt,
  aboveskip=0pt,
  belowskip=10pt,
  basicstyle=\ttfamily,
  showstringspaces=false,
  frame=ltrb,
  tabsize=4,
  numbers=left,
  numberstyle=\small,
  numbersep=8pt,
  morekeywords={*, factorial, sum, erlang},
  keywordstyle=\color{blue!70}, commentstyle=\color{red!50!green!50!blue!50},
}
\DeclareGraphicsExtensions{.eps,.ps,.jpg,.bmp}

\begin{document}

\title{Introduction to Analysis \\ Assignment 6}
\date{\today}
\author{Chuan Lu}

\maketitle

\problem{1}{Problem 29, Section 4.4, Page 89}
\solution{Sol}{
Both are not true. We can construct a $f$ like this:
$$
f(x) = \left\{
\begin{aligned}
1+\frac{1}{n^2}, ~n \le x < n+\frac{1}{2}, ~\forall n\in\mathbb{N} \\
-1, ~n+\frac{1}{2} \le x < n+1, ~\forall n\in\mathbb{N}
\end{aligned}
\right.
$$
Then $f$ is measurable, and $f$ is bounded on any bounded set, and
$$
a_n = \int_{n}^{n+1}f = \frac{1}{2n^2}.
$$
Clearly the series $\sum\limits_{n=1}^{\infty}a_n = \sum\limits_{n=1}^{\infty}\frac{1}{2n^2} $ converges absolutely, but 
$$
\int_{1}^{\infty}|f| = \sum_{n=1}^{\infty}(1+\frac{1}{2n^2}) = \infty,
$$
which means $f$ is not integratable on $[1, \infty)$.
}

\problem{2}{Problem 33, Section 4.4, Page 90}
\solution{Proof}{
First,
$$
|f_n-f| \le |f| + |f_n|, ~\forall n.
$$
Then since $f$ is integratable on $E$, if $\lim\limits_{n\to\infty}\int_E |f_n| = \int_{E}|f| $, we know $|f_n|+|f|$ converges pointwise a.e. to $2|f|$, and
$$
\lim_{n\to\infty}\int_E(|f_n|+|f|) = 2\int_E |f| < \infty,
$$
with General Lebesgue Dominated Convergence Theorem, notice $|f_n-f| $ converges pointwise a.e. to 0,
$$
\lim_{n\to\infty}\int_E|f_n-f| = \int_E 0 = 0.
$$
On the other hand, notice
$$
|f_n|-|f| \le |f_n-f|, ~\forall n.
$$
with the same method, since $\int_E|f-f_n|\to 0 $, and $|f-f_n| $ converges pointwise to 0,
$$
\lim_{n\to\infty}\int_E |f_n-f| = \int_E 0 = 0,
$$
we know from $|f_n|-|f| $ converges pointwise a.e. to 0,
$$
\lim_{n\to\infty}\int_E |f_n|-|f| = \int_E 0 = 0.
$$
Hence
$$
\lim_{n\to\infty}\int_E|f_n| = \int_E |f|.
$$
}

\end{document}