### 2017-11-30 Graduate - Introduction to Analysis I/Assignment 8/assignment.tex
add prob 1.1

%!TEX program = xelate
\input{structure.tex}
\usepackage{epstopdf}
\usepackage{graphics}
\usepackage{subfig}
\usepackage{listings}
\lstset{
  breaklines=true,
  xleftmargin=25pt,
  xrightmargin=25pt,
  aboveskip=0pt,
  belowskip=10pt,
  basicstyle=\ttfamily,
  showstringspaces=false,
  frame=ltrb,
  tabsize=4,
  numbers=left,
  numberstyle=\small,
  numbersep=8pt,
  morekeywords={*, factorial, sum, erlang},
  keywordstyle=\color{blue!70}, commentstyle=\color{red!50!green!50!blue!50},
}
\DeclareGraphicsExtensions{.eps,.ps,.jpg,.bmp}

\begin{document}

\title{Introduction to Analysis \\ Assignment 8}
\date{\today}
\author{Chuan Lu}

\maketitle

\problem{1}{Problem 37, Page 123}
\solution{Sol}{
(i) Let
$$
f(x) = \left\{
\begin{aligned}
&x\sin(\frac{1}{x}), ~0 < x \le 1, \\
&0, ~x = 0
\end{aligned}
\right.
$$
Then $f$ is continuous on $(0, 1]$. Since $\lim\limits_{x\to 0^+}f(x) = 0 = f(0)$, $f$ is continuous on $[0, 1]$. For $x_1, x_2 \in [\epsilon, 1]$ where $\epsilon > 0$, 
$$
\begin{aligned}
|f(x_1)-f(x_2)| &= |x_1\sin\frac{1}{x_1}-x_2\sin\frac{1}{x_2}| = |(x_1-x_2)\sin\frac{1}{x_1}+x_2(\sin\frac{1}{x_1}-\sin\frac{1}{x_2})| \\
&\le |x_1-x_2|\sin\frac{1}{x_1} + 2x_2|\cos\frac{1}{2}(\frac{1}{x_1}+\frac{1}{x_2})\sin\frac{1}{2}(\frac{1}{x_1}-\frac{1}{x_2})| \\
&\le |x_1-x_2|\sin\frac{1}{x_1} + 2x_2|\sin\frac{x_2-x_1}{2x_1x_2}| \le |x_1-x_2|\sin\frac{1}{x_1}+2x_2|\frac{x_2-x_1}{2x_1x_2}| \\
&= |x_1-x_2|(\sin\frac{1}{x_1}+\frac{1}{x_1}) \le |x_1-x_2|(1+\frac{1}{\epsilon}).
\end{aligned}
$$ 
Thus $f$ is Lipschitz, with Proposition 7 we know $f$ is absolutely continuous.

However, with Problem 35 we know $f$ is not of bounded variation on $[0, 1]$, and with Remark on Page 122 we know $f$ is not absolutely continuous on $[0, 1]$.
}

\problem{2}{Problem 39, Page 123}

\problem{3}{Problem 41, Page 123}

\problem{4}{Problem 49, Page 128}

\problem{5}{Problem 56, Page 129}

\problem{6}{Problem 59, Page 129}

\end{document}