### 2018-02-17 Graduate - Introduction to Analysis II/HW3/3HW20180204.tex
add prob 7, 8

\documentclass{article}%

\usepackage{amsmath}

\usepackage{graphicx}

\usepackage{amsfonts}%

\usepackage{amssymb}

\setlength{\topmargin}{-0.75in}

\setlength{\textheight}{9.25in}

\setlength{\oddsidemargin}{0.0in}

\setlength{\evensidemargin}{0.0in}

\setlength{\textwidth}{6.5in}

\def\labelenumi{\arabic{enumi}.}

\def\theenumi{\arabic{enumi}}

\def\labelenumii{(\alph{enumii})}

\def\theenumii{\alph{enumii}}

\def\p@enumii{\theenumi.}

\def\labelenumiii{\arabic{enumiii}.}

\def\theenumiii{\arabic{enumiii}}

\def\p@enumiii{(\theenumi)(\theenumii)}

\def\labelenumiv{\arabic{enumiv}.}

\def\theenumiv{\arabic{enumiv}}

\def\p@enumiv{\p@enumiii.\theenumiii}

\pagestyle{plain}

\setcounter{secnumdepth}{0}

\newtheorem{theorem}{Theorem}

\newtheorem{acknowledgement}[theorem]{Acknowledgement}

\newtheorem{algorithm}[theorem]{Algorithm}

\newtheorem{axiom}[theorem]{Axiom}

\newtheorem{case}[theorem]{Case}

\newtheorem{claim}[theorem]{Claim}

\newtheorem{conclusion}[theorem]{Conclusion}

\newtheorem{condition}[theorem]{Condition}

\newtheorem{conjecture}[theorem]{Conjecture}

\newtheorem{corollary}[theorem]{Corollary}

\newtheorem{criterion}[theorem]{Criterion}

\newtheorem{definition}[theorem]{Definition}

\newtheorem{example}[theorem]{Example}

\newtheorem{exercise}[theorem]{Exercise}

\newtheorem{lemma}[theorem]{Lemma}

\newtheorem{notation}[theorem]{Notation}

\newtheorem{problem}[theorem]{Problem}

\newtheorem{proposition}[theorem]{Proposition}

\newtheorem{remark}[theorem]{Remark}

\newtheorem{solution}[theorem]{Solution}

\newtheorem{summary}[theorem]{Summary}

\newenvironment{proof}[1][Proof]{\textbf{#1.} }{\ \rule{0.5em}{0.5em}}

\begin{document}

\begin{center}

\textbf{Homework 2}\bigskip

\end{center}

\noindent\textbf{Instructions}:References such as III.2.7 refer to Problem 7 in Section 2 of Chapter III in Conway's book.\smallskip

\noindent If you use results from books including Conway's, please be explicit about what results you are using.

\noindent \textbf{Reminder: Exam I will be from 6:30 to 10:30 on Thursday, February 15 in Room 40 of Schaeffer Hall.}

\begin{center}

\emph{Homework 3 is due in your Dropbox folder by 11:59, Sunday , February 19.}

\end{center}

\medskip

\noindent Working on this homework will help you with Exam I, so please don't put it off until after the exam.

\begin{enumerate}

\item Problem IV.2.4

\textbf{(a)}
By Abel's transform, let $\{a_n\}, \{b_n\}$ be two sequences, and $B_k = \sum\limits_{i=1}^{k}b_i $. Then
$$
\sum_{k=1}^{n}a_kb_k = a_nB_n - \sum_{k=1}^{n-1}(a_{k+1}-a_k)B_k.
$$
Hence for each fixed $n$, denote $\sum\limits_{k=1}^{n}a_k $ by $A_n $
$$
C_n = \lim_{r\to1^-}\sum_{k=1}^{n}a_kr^k = r^nA_n - \sum_{k=1}^{n-1}r^k(r-1)A_k.
$$
Since $\sum a_n(z-a)^n $ have radius of convergence 1,
$$
\lim_{r\to 1^-} \sum_{n=1}^{\infty}a_nr^n < \infty,
$$
then we can change the order of limits:
$$
\lim_{r\to 1^-}\lim_{n\to\infty}\sum_{k=1}^{n}a_kr^k = \lim_{n\to\infty}\lim_{r\to 1^-}\sum_{k=1}^{n}a_kr^k = \lim_{n\to\infty} A_n
$$
since each $A_k $ is a finite number, which comes from $\sum a_n $ converges to $A$. Hence,
$$
\lim_{r\to 1^-}\sum_{n=1}^{\infty}a_nr^n = \lim_{n\to\infty} A_n = A.
$$

\textbf{(b)}
Consider $a_n = \frac{(-1)^{n+1}}{n} $, then by Proposition III.1.4,
$$
\lim_{n\to\infty}\left|\frac{a_n}{a_{n+1}}\right| = \lim\frac{n+1}{n} = 1.
$$
Hence the series $\sum a_n(z-a)^n $ have radius of convergence 1. Since the series $\sum a_n $ is a Leibniz series, then it converges to $A < \infty$.

Now consider the function $f(z) = \log z$, it is analytic on $|z-1| < 1$, and it has power series expansion
$$
f(z) = \sum_{n=0}^{\infty}b_n (z-1)^n,
$$
where $b_n = \frac{1}{n!}f^{(n)}(1) = \frac{(-1)^{n-1}}{n} (n \ge 1), $ and $b_0 = 0 $. By this we can find $a_i = b_i $ for each $i \ge 0$, thus
$$
\sum a_n = f(2) = \log 2
$$

\item Problem IV.2.6

\textbf{Sol.} In the region where $f(z) = \sqrt{z}$ is analytic, 
$$
f(z) = \sum_{n=0}^{\infty}a_n(z-1)^n,
$$
where
$$
a_n = \frac{1}{n!}f^{(n)}(1) = \frac{(-1)^{(n-1)}}{n!}\frac{(2n-3)!!}{2^{n}} (n \ge 1), ~a_0 = 1.
$$
and since
$$
\lim_{n\to\infty}\left|\frac{a_n}{a_{n+1}}\right| = \lim_{n\to\infty} \frac{2n+2}{2n-1} = 1,
$$
we know the radius of convergence is 1.

\item Problem IV.2.9

\textbf{(a)}
Let $f(z) = e^{z}-e^{-z} $, then by Corollary 2.13,
$$
f^{(n-1)}(0) = \frac{(n-1)!}{2\pi i}\int_{\gamma}\frac{e^z-e^{-z}}{z^n}dz = 1+(-1)^{n}.
$$
Thus
$$
\int_{\gamma}\frac{e^z-e^{-z}}{z^n}dz = \frac{2\pi i}{(n-1)!}(1+(-1)^n).
$$

\textbf{(b)}
Let $f(z) = 1$, then 
$$
f^{(n-1)}(\frac{1}{2}) = \frac{(n-1)!}{2\pi i}\int_{\gamma}\frac{1}{(z-\frac{1}{2})^n} dz = \left\{
\begin{aligned}
& 1, n = 1 \\
& 0, n \ge 2
\end{aligned}
\right.
$$
Thus
$$
\int_{\gamma}\frac{1}{(z-\frac{1}{2})^n} dz = \left\{
\begin{aligned}
& 2\pi i, n = 1 \\
& 0, n \ge 2
\end{aligned}
\right.
$$

\textbf{(c)}
First, we have
$$
\frac{1}{z^2+1} = \frac{1}{2i}(\frac{1}{z-i}-\frac{1}{z+i}).
$$
Then let $f(z) = 1 = g(z)$, then
$$
1 = f(i) = \frac{1}{2\pi i}\int_{\gamma}\frac{1}{z-i}dz = g(-i) = \frac{1}{2\pi i}\int_{\gamma}\frac{1}{z+i}dz.
$$
Hence 
$$
\int_{\gamma}\frac{1}{z^2+1}dz = \frac{1}{2i}(\int_\gamma\frac{dz}{z-i}-\int_\gamma\frac{dz}{z+i}) = 0.
$$

\textbf{(d)}
Let $f(z) = \sin z$, then $f$ is analytic on $\mathbb{C}$.
$$
f(0) = \frac{1}{2\pi i}\int_{\gamma}\frac{\sin z}{z}dz = 0.
$$
Hence
$$
\int_\gamma \frac{\sin z}{z}dz = 0.
$$

\textbf{(e)}
Let $f(z) = z^{1/m} $, then
$$
f^{(m-1)}(1) = \frac{(m-1)!}{2\pi i}\int_{\gamma}\frac{z^{1/m}}{(z-1)^m}dz = \prod_{i=0}^{m-1}(\frac{1}{m}-i).
$$
Hence
$$
\int_{\gamma}\frac{z^{1/m}}{(z-1)^m}dz  = \frac{2\pi i}{(m-1)!}\prod_{i=0}^{m-1}(\frac{1}{m}-i).
$$

\item Problem IV.2.11

First, 
$$
f(z) = \frac{1}{2i}\log(\frac{1+iz}{1-iz}) = \frac{1}{2i}(\log(1+iz)-\log(1-iz)).
$$
We know $\log(z)$ is analytic on $\mathbb{C}\setminus l$, where $l$ is a line starting from the origin. Thus $f$ is analytic on $\mathbb{C}\setminus l\{z_1, z_2\}$, where $z_1 = -i, ~z_2 = i $. Hence $l: Re(z) = 0 $.

On a branch of $f$, 
$$
\tan f(z) = \frac{1}{i}\frac{e^{if(z)}-e^{-if(z)}}{e^{if(z)}+e^{-if(z)}}.
$$
Since
$$
e^{if(z)} = e^{\frac{1}{2}\log(\frac{1+iz}{1-iz})} = (\frac{1+iz}{1-iz})^{1/2},
$$
we have
$$
\tan f(z) = \frac{1}{i}\frac{(\frac{1+iz}{1-iz})^{1/2}-(\frac{1+iz}{1-iz})^{-1/2}}{(\frac{1+iz}{1-iz})^{1/2}+(\frac{1+iz}{1-iz})^{-1/2}} = \frac{1}{i}\frac{2iz}{2} = z.
$$
When $|z| < 1$, $f(z)$ is analytic. Hence by Theorem 2.8, 
$$
\begin{aligned}
\log(1+iz) - \log(1-iz) &= \sum_{n=1}^{\infty}(-1)^{n-1}\frac{((1+iz)-1)^n}{n}-\sum_{n=1}^{\infty}(-1)^{n-1}\frac{((1-iz)-1)^n}{n} \\
&= \sum_{n=1}^{\infty}(-1)^{n-1}\frac{(iz)^n}{n}-\sum_{n=1}^{\infty}(-1)^{n-1}\frac{(-iz)^n}{n} = 2i\sum_{n=0}^{\infty}(-1)^n\frac{z^{2n+1}}{2n+1}.
\end{aligned}
$$
Hence $f(z) = \sum\limits_{n=0}^{\infty}(-1)^n\frac{z^{2n+1}}{2n+1} $.

\item Problem IV.3.3

\textbf{Sol.}
The only function is $f(z) = \exp(z)$.

First, it is clear that $f(z) = \exp(z)$ is a solution. Now suppose there is another function $g$, s.t. $g(x) = \exp(x), \forall x\in\mathbb{R}$, then for each point $x_0\in\mathbb{R} $, let $\{x_n\} = \{x_0 + \frac{1}{n}\}$, then $f(x_i) = g(x_i)$ for each $i$, and $\{x_n\}\to x_0 $. Hence, $x_0 $ is a limit point of the set $\{f = g\}$, thus by Corollary 3.8, $f = g$ on $\mathbb{C}$.

\item Problem IV.3.6

\textbf{Proof.} If $f(a) = 0$, then the proposition holds. Now suppose $f(a) \ne 0$, hence $|f(a)| > 0$. Let $g(z) = \frac{1}{f(z)}$, then since $f$ is analytic, $g$ is analytic in $G$. Hence $|g(a)| = \left|\frac{1}{f(a)}\right| = \frac{1}{|f(a)|} \ge |g(z)|$ for each $z\in G$. Then by Maximum Modulus Thm, $g$ is constant, which means $f$ is constant.

\item Problem IV.4.2

\textbf{Sol.} Checked https://math.stackexchange.com/questions/2422165/finding-a-closed-rectifiable-curve-with-prescribed-winding-numbers/2422214

Let $\gamma$: $[0, 1] \to \mathbb{C}$ be a union of circles:
$$
\gamma([\frac{1}{2}-\frac{1}{2^n}, \frac{1}{2}-\frac{1}{2^{n+1}}]) = \gamma_1 = B(\frac{1}{2^n}, \frac{1}{2^n}), ~n\ge 1,
$$
and
$$
\gamma([1-\frac{1}{2^n}, 1-\frac{1}{2^{n+1}}]) = \gamma_2 = B(-\frac{1}{2^n}, \frac{1}{2^n}), n \ge 1.
$$
And the direction of $\gamma_1 $ is counterclockwise and $\gamma_2 $ is clockwise.

Then $\gamma$ is rectifiable since length of $\gamma$ is $2\sum_{n=1}^{\infty}2\pi\frac{1}{2^n} = 4\pi$.
Now we show that $\gamma$ is continuous. In fact, we only need to show that $\gamma$ is continuous at the point $t = \frac{1}{2}$ and $t = 1$ since circles are continuous. First we know $\lim_{t\to 1}\gamma(t) = 0 $, hence $\gamma $ is closed. For $\forall \epsilon > 0$, pick $n$, s.t. $\frac{1}{2^n} < \frac{\epsilon}{2}$, then $d(\gamma(t), 0) < 2\times\frac{1}{2^n} < \epsilon$ for all $t\in [1-\frac{1}{2^n}, 1]$. Hence $\gamma$ is continuous at $t = 1$, and similarly, $\gamma$ is continuous at $t = \frac{1}{2}$.

For any integer $k$, if $k > 0$, pick a point $a$ inside the $k^{th} $ largest circle and outside the $(k+1)^{th} $ circle of $\gamma_1 $, then by Proposition 4.1, $n(\gamma; a) = k$. If $k = 0$, pick $a$ outside the largest circle of both $\gamma_1 $ and $\gamma_2 $, then $n(\gamma; a) = 0$.
 If $k < 0$, then pick $a$ inside the $k^{th} $ largest circle and outside the $(k+1)^{th} $ circle of $\gamma_2 $, by Prop 4.1 and Prop 4.3 $n(\gamma; a) = k$.

\item Problem IV.4.3

\textbf{Proof.} By Corollary 3.6, since $p$ is a polynomial of degree $n$, we know
$$
p(z) = c(z-z_1)(z-z_2)\cdots(z-z_n),
$$
where $z_i $ are roots of $p$ and $z_i $ could equal to $z_j $. Then
$$
\frac{p'(z)}{p(z)} = \frac{a\sum_{i=1}^{n}\prod_{j\ne i}^{n}(z-z_j)}{a\prod_{i=1}^{n}(z-z_j)} = \sum_{i=1}^{n}\frac{1}{z-z_i},
$$
with $z_i \in B(0, R) $.
Hence,
$$
\int_{\gamma}\frac{p'(z)}{p(z)}dz = \int_{\gamma}\sum_{i=1}^{n}\frac{1}{z-z_i}dz = \sum_{i=1}^{n}
\int_\gamma\frac{1}{z-z_i}dz = \sum_{i=1}^{n}2\pi i = 2\pi in
$$
by Prop 4.3 and Thm 4.4.

\item Problem IV.5.7

\item Problem IV.5.9

\end{enumerate}

\end{document}